Level-Set method on \texttt{p4est}

### 2015-06-04 paper/tex/abstract.tex
adding placeholder for carsten

%!TEX root = draft.tex
\begin{abstract}
We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids. The algorithms are based on a domain decomposition technique and implemented using \texttt{MPI} and the open-source \texttt{p4est} library. An important contribution is a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions. This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids. Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation. Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores currently. Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detail that would be impossible to achieve without a parallel adaptive strategy. We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
\end{abstract}
\begin{keyword}
Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
\end{keyword}
\maketitle

>>>>

%!TEX root = draft.tex
\begin{abstract}
We present scalable parallel algorithms for the level-set method on adaptive Quadtree and Octree Cartesian grids. The algorithms are based on a \textcolor{red}{domain decomposition technique} and implemented using \texttt{MPI} and the open-source \texttt{p4est} library. We present a scalable parallel semi-Lagrangian method which, similar to its serial implementation, is free of any time-step restrictions. This is achieved by introducing a scalable global interpolation scheme on adaptive tree-based grids. Moreover, we present a simple parallel reinitialization scheme using the pseudo-time transient formulation. Both parallel algorithms scale on the Stampede supercomputer, where we are limited to 4096 cores currently. Finally a relevant application of the algorithms is presented in modeling the crystallization phenomena by solving a Stefan problem, illustrating a level of detail that would be impossible to achieve without a parallel adaptive strategy. We believe that the algorithms presented in this article will be of interest and useful to researchers working with the level-set framework and modeling multi-scale physics in general.
\end{abstract}
\begin{keyword}
Quadtree/Octree Grids \sep Parallel Computing \sep Space Filling Curves \sep Semi-Lagrangian Method \sep Level-set Method
\end{keyword}
\maketitle

### 2015-06-04 paper/tex/introduction.tex
adding placeholder for carsten

%!TEX root = draft.tex
\section{Introduction}\label{sec:introduction}
The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes. As a result, the level-set method has been used to a wide range of applications such as multiphase flows, image segmentation, and computer graphics \cite{Osher;Fedkiw:02:Level-Set-Methods-an,Sethian:99:Level-set-methods-an}. An important feature of this method is that the location of the interface is defined implicitly on an underlying grid. This convenience, however, comes at a price. First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem, although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. Second, the level-set function has to be defined in a higher dimensional space than that of the interface. If only the location of the interface is needed, the added dimension greatly increases the overall computational cost. One way to avoid this problem is by computing the level-set only close to the interface, e.g.\ as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.

@@

The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been used by many researchers. For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}. \texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an Octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}. More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid. This forest is then partitioned in parallel using a global Morton curve. The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}. In fact the algorithms presented in this paper are implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}. We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.

>>>>

%!TEX root = draft.tex
\section{Introduction}\label{sec:introduction}
The level-set method, originally proposed by Sethian and Osher \cite{Osher;Sethian:88:Fronts-Propagating-w}, is a popular and powerful framework for tracking arbitrary interfaces that undergo complicated topological changes. As a result, the level-set method has been used to a wide range of applications such as multiphase flows, image segmentation, and computer graphics \cite{Osher;Fedkiw:02:Level-Set-Methods-an,Sethian:99:Level-set-methods-an}. An important feature of this method is that the location of the interface is defined implicitly on an underlying grid. This convenience, however, comes at a price. First, compared to an explicit method, e.g.\ front tracking \cite{Juric:96:A-Front-Tracking-Met, Tryggvason;Bunner;Esmaeeli;etal:01:A-Front-Tracking-Met}, the level-set method is typically less accurate and mass conservation could be a problem, although progress has been made in resolving this issue \cite{Enright;Fedkiw;Ferziger;etal:02:A-Hybrid-Particle-Le}. Second, the level-set function has to be defined in a one dimension higher space than that of the interface. If only the location of the interface is needed, the added dimension greatly increases the overall computational cost for uniform grids. One way to avoid this problem is by computing the level-set only close to the interface, e.g.\ as in the narrow-band level-set method \cite{Adalsteinsson;Sethian:95:A-Fast-Level-Set-Met} or, more recently, by using a hash table to restrict both computation and storage requirements \cite{Brun;Guittet;Gibou:12:A-local-level-set-me}.

@@

The idea of using SFCs for parallel partitioning of Quad-/Octrees is not new in itself and has been used by many researchers. For instance, \texttt{Octor} \cite{Tu;OHallaron;Ghattas:05:Scalable-parallel-oc} uses a Morton curve (also known as Z-curve) for traversing the leaves of an Octree for indexing and load balancing and has been scaled up to 62,000 cores \cite{Burstedde;Ghattas;Gurnis;etal:08:Scalable-adaptive-ma}. \texttt{Dendro} \cite{Sampath;Adavani;Sundar;etal:08:Dendro:-parallel-alg} is another example of an Octree code in which similar ideas are used for parallel partitioning and development of a parallel geometric multigrid that has been scaled up to about 32,000 cores \cite{Sampath;Biros:10:A-parallel-geometric}. More recently, authors in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo} extended these ideas to a collection, or a ``forest'', of Octrees that are connected through a common, potentially unstructured, hexahedral grid. \textcolor{red}{How about adding the p4est comments here ?} This forest is then partitioned in parallel using a global Morton curve. The implementation of these algorithms, which were shown to scale to more than 200,000 cores, is publicly available through a simple API provided by the \texttt{p4est} library \cite{p4est-github}. In fact the algorithms presented in this paper are implemented on top of the \texttt{p4est} library and we do not discuss any algorithm that is already covered in \cite{Burstedde;Wilcox;Ghattas:11:p4est:-Scalable-Algo}. We also make use of the popular \texttt{PETSc} \cite{Balay;Abhyankar;Adams;etal:14:PETSc-Web-page} library for linear algebra and its parallel primitives, such as parallel ghosted vector and scatter/gather operations, which simplifies the implementation.
